---
title: "2.8 Exploratory Data Analysis with R"
author: "T.A. Meraxa"
date: "October 24, 2015"
output: html_document
---

#### 1. Consider the data CPS 1985 from Berndt (1991)

```{r results='hide', message=FALSE, warning=FALSE}
# Load library (AER)
library(AER)

# Load data
data("CPS1985")
```

```{r}
# see the data structure and summary
str(CPS1985)
summary(CPS1985)
```


#### 1-1. One numerical variable

```{r}
# Attach the variable names
attach(CPS1985)

# Now let's look at a numerical variable
summary(wage)
median(wage)
var(wage)
sd(wage)
```


#### 1-2. Graphical summary

```{r}
# Histogram
hist(wage, freq = FALSE)
hist(log(wage), freq = FALSE)
lines(density(log(wage)),col = 4)
```


#### 1-3. One categorical variable
For this type of variable, R categorizes it as factors. It is useless to calculate the mean and variance.

```{r}
# Now let's look at a categorical variable
summary(occupation)

# Proportion tables can also be used
tab <- table(occupation)
prop.table(tab)

# Categorical variables are best shown using barplots
barplot(tab)
pie (tab)
```


#### 1-4. Two categorical variables
Two categorical variables can be summarized using contingency table, using 'xtabs()', or simply 'table()'
```{r}
# Now let's look at a categorical variable
xtabs(~ gender + occupation, CPS1985)

# plot it using boxplot
plot(gender ~ occupation)
```


#### 1-5. Two numerical variables
Exemplify exploratory analysis on **wage** and **education**. To examine the relationship use `cor()`.
```{r}
# Examine correlation
cor(log(wage),education)
cor(log(wage),education, method = "spearman")

# PLot the correlation
plot(log(wage) ~ education)
```

#### 1-6.
It is common for data frame to have both numerical and categorical example
```{r}
# Function for numerical summaries
tapply(log(wage),gender,mean)

# Plot the summaries
plot(log(wage) ~ gender)

#plot it using QQplot
mwage <- subset(CPS1985, gender == "male")$wage
fwage <- subset(CPS1985, gender =="female")$wage
qqplot(mwage, fwage, 
       xlim = range(wage), ylim = range(wage), 
       xaxs = "i", yaxs = "i", 
       xlab = "male", ylab = "female")
abline(0,1)
```
