---
title: "3_Linear_Regression"
author: "T.A. Meraxa"
date: "July 3, 2016"
output: html_document
---

### 3. Linear Regression

In R models are fitted using model-fitting function, that is 'lm()'. Usage is as follows 
'fm <- lm(formula,data,..)'. Generic functions such as `summary()`,`residuals()`,or `predict()` are available for all standard models.

#### 3.1. Simple Linear Regression

```{r}
# Load data
data("Journals")
journals <- Journals[,c("subs","price")]
Journals$citeprice <- Journals$price/Journals$citations
```

```{r}
# see the data structure and summary
summary(Journals)
```

The goal is to estimate the effect of the price per citation on the number of library subscription. Therefore, we fit a linear regression model

log(subs)[i]= Beta[1] + Beta[2]log(citeprice)+Epsilon[i]

The primary function of tthe `lm()` estimates a linear regression using ordinary least squares

```{r}
# plot the regression fit
plot(log(subs) ~ log(citeprice), data = journals)
jour_lm <- lm(log(subs) ~ log(citeprice), data = journals)
abline(jour_lm)
```
