---
title: "3_Linear_Regression"
author: "T.A. Meraxa"
date: "July 3, 2016"
output: html_document
---

### 3. Linear Regression

In R models are fitted using model-fitting function, that is `lm()`. Usage is as follows 
`fm <- lm(formula,data,..)`. Generic functions such as `summary()`,`residuals()`,or `predict()` are available for all standard models.

#### 3.1. Simple Linear Regression

```{r}
# Load data
data("Journals")
journals <- Journals[,c("subs","price")]
Journals$citeprice <- Journals$price/Journals$citations
```

```{r}
# see the data structure and summary
summary(Journals)
```

The goal is to estimate the effect of the price per citation on the number of library subscription. Therefore, we fit a linear regression model

log(subs)[i]= Beta[1] + Beta[2]log(citeprice)+Epsilon[i]

The primary function of tthe `lm()` estimates a linear regression using ordinary least squares

```{r}
# plot the regression fit
plot(log(subs) ~ log(citeprice), data = journals)
jour_lm <- lm(log(subs) ~ log(citeprice), data = journals)
abline(jour_lm)
```

```{r}
# check the class of fitted model jour_lm
class(jour_lm)

# check the names of its components
names(jour_lm)

# check the summary of jour_lm
summary(jour_lm)

# create the summary for the fitten lm object
jour_slm <- summary(jour_lm)
class(jour_slm)
names (jour_slm)

#check the coefficient of jour_slm
jour_slm$coefficients
```

**Useful functions**

* `print()`     simple printed display
* `summary()`   standard regression output
* `coef()`      (or `coefficients()`) extracting the regression coefficients
* `residuals()` (or `resid()`) extracting residuals
* `fitted()`    (or `fitted.values()`) extracting fitted values
* `anova()`     comparison of nested models
* `predict()`   predictions for new data
* `plot()`      diagnostic plots
* `confint()`   confidence intervals for the regression coefficients
* `deviance()`  residual sum of squares
* `vcov()`      (estimated) variance-covariance matrix
* `logLik()`    log-likelihood (assuming normally distributed errors)
* `AIC()`       information criteria including AIC, BIC/SBC (assuming normally distributed errors)

##### Analysis of Variance

```{r}
# perform analysis of variance
anova(jour_lm)
```

##### Point and interval estimates

```{r}
# extracting the estimated regression coefficients
coef(jour_lm)

# obtain confidence interval
confint(jour_lm, level = 0.95)
```

##### prediction

```{r}
# predicting points on the regression line and predicting a new data value
predict(jour_lm, newdata = data.frame(citeprice = 2.11), interval = "confidence")
predict(jour_lm, newdata = data.frame(citeprice = 2.11), interval = "prediction")
        
#
lciteprice <- seq(from = -6, to = 4, by = 0.25)
jour_pred <- predict(jour_lm, interval = "prediction",
                     newdata = data.frame(citeprice = exp(lciteprice)))
```

##### plotting "lm" objects

```{r}
# predicting points on the regression line and predicting a new data value
par(mfrow = c(2,2))
plot(jour_lm)
par(mfrow = c(1,1))
```

##### testing linear hypothesis

```{r}
# linear hypothesis test
linearHypothesis(jour_lm, "log(citeprice) = -0.5")
```

#### 3.2. Multiple Linear Regression
