---
title: "3_Linear_Regression_Part_2"
author: "T.A. Meraxa"
date: "August 20, 2016"
output: html_document
---

```{r}
# data prep
library(AER)
library(splines)
data("CPS1988")
attach(CPS1988)
cps_lm <- lm(log(wage) ~ experience + I(experience^2) + education + ethnicity, data = CPS1988)
```

### 3. Linear Regression (continued)

#### 3.3 Partly Linear Model

Quadrati terms in experience are common in wage equations; however, given the size of the CPS1988 data, it may be worthwile to model the role of thisvariable using more flexible tools.

Consider the model:

log(wage)= Beta[1] + g(experience) + Beta[2]education + Beta[3]ethnicity + Epsilon

Here, g is an unknown function to be estimated from the data. Regression splines are use for this task. Splines is included in the packages `("splines")` and also in `("AER")`. 

```{r}
# computing B splines
cps_plm <- lm(log(wage) ~ bs(experience, df = 5) + education + ethnicity, data = CPS1988)
summary(cps_plm)
```

The choice of `df = 5` is based on Schwarz criterion (BIC). See the code below.

```{r}
# computing B splines; Schwarz criterion (BIC) selection 
cps_bs <- lapply(3:10, function (i) lm(log(wage) ~
          bs(experience, df = i) + education + ethnicity,
          data = CPS1988))

structure(sapply(cps_bs, AIC, k = log(nrow(CPS1988))),
          .Names = 3:10)
```

The cubic spline selected from the model is comparred with the classical fit form `cps_lm`

```{r}
# plot a graph
cps <- data.frame(
    experience = -2:60, 
    education = with (CPS1988, mean(education[ethnicity == "cauc"])),
    ethnicity = "cauc")

cps$yhat1 <- predict(cps_lm, newdata = cps)
cps$yhat2 <- predict(cps_plm,newdata = cps)
plot(log(wage) ~ jitter(experience, factor = 3), 
     pch = 19,
     col = rgb(0.5, 0.5, 0.5, alpha = 0.02), 
     data = CPS1988)
lines(yhat1 ~ experience, data = cps, lty = 2)
lines(yhat2 ~ experience, data = cps)
legend("topleft", c("quadratic", "spline"), lty = c(2,1), bty = "n")

```

The spline are not too distinct for the 20-40 years of experience range. Overall, the spline version exhibits less curvature beyond eight years of experience. The highlight of this feature is that the more pronounced curvature below seven years of experience.

### 4. Factors, Interactions and Weights

Formula       | Description
------------  |------------
y ~ a + x     | Model without interaction: indentical slopes with respect to x, but different intercepts with respect to a.
Y ~ a* x      | Model with interaction: the term a:x gives the difference in slopes compared with the reference category.
y ~ a/x       | Model with interaction: produces the same fitten values as the mdoel above but using a nested coefficient coding. An explicit slope estimate is computed for each category in a.
y ~ (a+b+c)^2 | Model with all two-way interactions(excluding three way interaction).

#### Interactions

```{r}
# Let's consider an interaction between `ethnicity` and `education`
cps_int <- lm(log(wage) ~ experience + I(experience^2) + education * ethnicity, data = CPS1988)
coeftest(cps_int)
```

Resuts indicated that the interaction term is significant at the 5% level